Investigating Predictors for American Attitudes on Justifiability of Abortion

### 2022-06-20 Final/FinalReport.qmd
minor edits to introduction

The dataset used in this analysis observes attitudes on the
justifiability of abortion among respondents in in the United States
across six "waves" of the World Values Survey (1982-2011), which is
administered every few years and collects information about people's
values and beliefs worldwide, alongside basic demographic
characteristics. The data are collected via face-to-face interviews at
the respondents' homes. In this data set, the response variable,
justifiability of abortion, is a numerical measure on a scale of 1 to 10
on the individual person's attitude toward whether abortion is
justifiable or not. Individuals responded 1 for "abortion is never
justified" and 10 for "abortion is always justified.

Our research question is as follows: "Do an individual's political
ideology and their personal attitudes/preferences towards other issues,
such as the importance of religion in their life and their respect for
authority, among others, predict their attitude on the justifiability of
abortion?" We will attempt to answer this question using an EDA-informed
predictive model. Given that this issue has become highly polarized by
political party, we predict that liberal ideology and liberal-leaning
attitudes on other issues will correlate with belief that abortion is
more justified.

>>>>

The dataset used in this analysis observes attitudes on the justifiability of abortion among respondents in in the United States across six “waves” of the World Values Survey (1982-2011), which is administered every few years and collects information about people’s values and beliefs worldwide, alongside basic demographic characteristics. The data are a nationally representative sample of the American population and are collected via face-to-face interviews at the respondents’ homes. In this data set, the response variable, justifiability of abortion, is a numerical measure on a scale of 1 to 10 on the individual person’s attitude toward whether abortion is justifiable or not.  Individuals responded 1 for “abortion is never justified” and 10 for “abortion is always justified.

Our research question is as follows: “Do an individual’s political ideology and their personal attitudes/preferences towards other issues, such as the importance of religion in their life and their respect for authority, among others, predict their attitude on the justifiability of abortion?” We will attempt to answer this question using an EDA-informed predictive model. Because the data are nationally representative, we can infer that all the observations are independent and the variables of interest may show a linear relationship. Given that this issue has become highly polarized by political party, we predict that liberal ideology and liberal-leaning attitudes on other issues will correlate with belief that abortion is more justified.
